Capturing the geometric structure of our experiences and how we remember them

### 2018-08-16 paper/main.tex
what if we shelved the fmri analysis

\subsubsection{Brain regions that mirror the correlational structure of naturalistic memories}.
While each participant remembered roughly the same sequence of events, there were subtle idiosyncracies in the way each individual described the events. For example, the event by event similarity structure (e.g. the off-diagonal values of the recall correlation matrices) differed across individuals (Supp. Fig.~\ref{fig:corrmats}). We hypothesized that idiosyncracies in the recall dynamics across subjects might be related to the temporal structure of brain networks are representing the events during video viewing. Thus in the last analysis, we identified brain areas whose responses at the time of viewing exhibited the same (segmented) idiosyncratic dynamics that later appeared in that participants’ recalls.

The amount of time/number of sentences used describing each scene differed drastically between participants and as compared to the original video \cite{ChenEtal17}. To account for these temporal non-linearities between the video and recall, we used dynamic time warping to align each participant's recall model and the movie model (DTW REF).  We used the resulting alignment path to warp the fMRI timeseries (for each voxel) as well as the recall model timeseries.  Then using a searchlight procedure (5mm), we computed the timepoint-by-timepoint neural correlation matrix within each searchlight. To measure the similarity between the neural and behavioral correlation matrices, we vectorized and then correlated them. The result was a correlation value for each voxel representing the extent to which the brain data mirrored the recall data.  Finally, to test the significance of the correlation values, we phase shifted the recall models (100 permuations) and recomputed the correlation values. Significance was determined by ``real'' correlation values that exceeded the 95th percentile of this null distribution. The resulting map suggests that the correlational structure of a network of brain regions including the ventromedial prefrontal cortex, anterior cingulate, and right medial temporal lobe mirror the idiosyncratic structure of participants'' recalls.

\begin{figure}[t!]
\centering
\includegraphics[width=1\textwidth]{figs/6_brains.pdf}
\caption{\small \textbf{Recall model correlational structure searchlight analysis.} A). We used dynamic time warping to separately align each participant's recall model to the video model. The left matrix is S13's video-recall correlation matrix before warping. The yellow stars represent the path recovered by dynamic time warping. The right matrix is the result after warping. The same warping path was applied to the fMRI timeseries (separately for each participant). Below the matrices are toy examples representing a vectorized (aligned) recall correlation matrix and a vectorized (aligned) fMRI searchlight correlation matrix. These vectors were correlated for each searchlight sphere in the brain to assess brain regions whose temporal correlation structure mirrored the participant's recall model correlation structure. B) A surface plot of brain regions whose activity patterns mirrored the participant-specific recall correlation matrix. Significance threshold (p$<$.05) derived from a permutation procedure by phase shifting the video model and recomputing the z-scores.}
\label{fig:brainz}
\end{figure}

@@

Prior work on neural responses during naturalistic experience and recall (e.g. Hasson group) has largely focused on identifying brain regions whose responses are reliably similar across individuals. This allows one to identify which regions might be processing or representing the stimulus, or retrieving details of the experience during recall. Our approach is fundamentally different. We ask: which brain regions during movie viewing match how individual participants will recall the movie later (in terms of the temporal correlations of their recall topic trajectory)? This approach highlights regions such as the vmPFC that may respond in idiosyncratic ways across individuals, but that nonetheless play an important role in encoding experience into memory.

@@

In summary, using a novel content-based memory assessment approach,

+ novel way to quantify naturalistic experiences and memories
+ opens the door to studies of how various manipulations (or simply the passage of time) might lead to changes in the structure of naturalistic memories
+ other potential applications - quantifying learning from educational videos, memory assessment/diagnosis of memory disorders and other psychiatric disorders

>>>>

%NOTE: I'm not sure the brain stuff really fits in this paper.  It's cool, but seems like a distraction from the main point.  perhaps it needs it own paper with more motivation, and more extensive analyses (for example, what brain regions have temporal structure like the movie model?). I'm also having a hard time motivating it in the discussion, so maybe we should cut it.

% \subsubsection{Brain regions that mirror the correlational structure of naturalistic memories}.
% While each participant remembered roughly the same sequence of events, there were subtle idiosyncracies in the way each individual described the events. For example, the event by event similarity structure (e.g. the off-diagonal values of the recall correlation matrices) differed across individuals (Supp. Fig.~\ref{fig:corrmats}). We hypothesized that idiosyncracies in the recall dynamics across subjects might be related to the temporal structure of brain networks are representing the events during video viewing. Thus in the last analysis, we identified brain areas whose responses at the time of viewing exhibited the same (segmented) idiosyncratic dynamics that later appeared in that participants’ recalls.
%
% The amount of time/number of sentences used describing each scene differed drastically between participants and as compared to the original video \cite{ChenEtal17}. To account for these temporal non-linearities between the video and recall, we used dynamic time warping to align each participant's recall model and the movie model (DTW REF).  We used the resulting alignment path to warp the fMRI timeseries (for each voxel) as well as the recall model timeseries.  Then using a searchlight procedure (5mm), we computed the timepoint-by-timepoint neural correlation matrix within each searchlight. To measure the similarity between the neural and behavioral correlation matrices, we vectorized and then correlated them. The result was a correlation value for each voxel representing the extent to which the brain data mirrored the recall data.  Finally, to test the significance of the correlation values, we phase shifted the recall models (100 permuations) and recomputed the correlation values. Significance was determined by ``real'' correlation values that exceeded the 95th percentile of this null distribution. The resulting map suggests that the correlational structure of a network of brain regions including the ventromedial prefrontal cortex, anterior cingulate, and right medial temporal lobe mirror the idiosyncratic structure of participants'' recalls.
%
% \begin{figure}[t!]
% \centering
% \includegraphics[width=1\textwidth]{figs/6_brains.pdf}
% \caption{\small \textbf{Recall model correlational structure searchlight analysis.} A). We used dynamic time warping to separately align each participant's recall model to the video model. The left matrix is S13's video-recall correlation matrix before warping. The yellow stars represent the path recovered by dynamic time warping. The right matrix is the result after warping. The same warping path was applied to the fMRI timeseries (separately for each participant). Below the matrices are toy examples representing a vectorized (aligned) recall correlation matrix and a vectorized (aligned) fMRI searchlight correlation matrix. These vectors were correlated for each searchlight sphere in the brain to assess brain regions whose temporal correlation structure mirrored the participant's recall model correlation structure. B) A surface plot of brain regions whose activity patterns mirrored the participant-specific recall correlation matrix. Significance threshold (p$<$.05) derived from a permutation procedure by phase shifting the video model and recomputing the z-scores.}
% \label{fig:brainz}
% \end{figure}

@@

% Prior work on neural responses during naturalistic experience and recall (e.g. Hasson group) has largely focused on identifying brain regions whose responses are reliably similar across individuals. This allows one to identify which regions might be processing or representing the stimulus, or retrieving details of the experience during recall. Our approach is fundamentally different. We ask: which brain regions during movie viewing match how individual participants will recall the movie later (in terms of the temporal correlations of their recall topic trajectory)? This approach highlights regions such as the vmPFC that may respond in idiosyncratic ways across individuals, but that nonetheless play an important role in encoding experience into memory.

@@
